---
title: "Tools and Workflows for Reproducible Research in the Quantitative Social Sciences"
subtitle: "Advanced R Markdown & LaTeX"
author: "Bernd Weiß<br /><b><i>Johannes Breuer</b></i><br />Arnim Bleier"
date: "2021-11-18"
location: "GESIS, online"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["default", "default-fonts", "workshop.css"]
    nature:
      highlightStyle: "github"
      highlightLines: true
      countIncrementalSlides: false
---

layout: true

```{r setup, include = F}
if (!require(easypackages)) install.packages("easypackages")
library(easypackages)

packages("knitr",
         "rmarkdown",
         "gadenbuie/xaringanExtra",
         "gadenbuie/tweetrmd",
         "hadley/emo",
         prompt = F)

options(htmltools.dir.version = FALSE,
        htmltools.preserve.raw = FALSE)

opts_chunk$set(echo = TRUE,
               fig.align = "center")

xaringanExtra::use_xaringan_extra(c("tile_view", "clipboard"))
xaringanExtra::use_extra_styles(hover_code_line = TRUE,
                                mute_unhighlighted_code = FALSE)
```

<div class="my-footer">
  <div style="float: left;"><span>`r gsub("<br />", ", ", gsub("<br /><br />|<a.+$", "", metadata$author))`</span></div>
  <div style="float: right;"><span>`r metadata$location`, `r metadata$date`</span></div>
  <div style="text-align: center;"><span>`r gsub(".+<br />", " ", metadata$subtitle)`</span></div>
</div>

---

# Content of this session

In the previous session on `R Markdown`, we covered some of its basics and focused on `HTML` output. In this session, we will cover some more advanced topics and focus on generating PDF output using $\LaTeX$. In addition to combining some of the things we have discussed on the previous session on $\LaTeX$ with `R Markdown`, we will mainly deal with two topics:

- More on the creation and customization of tables in `R Markdown`

- Bibliographies and citations

---

## PDF output with `R Markdown`

If you want to generate PDF output with `R Markdown`, you need $\LaTeX$. If you have a $\LaTeX$ distribution like [`MiKTeX`](https://miktex.org/) or [`TeX Live`](https://www.tug.org/texlive/) on your system, you should be all set.

If you do not have $\LaTeX$ installed, the easiest option - especially if you do not want to use plain $\LaTeX$ - is installing [`TinyTeX`](https://yihui.org/tinytex/), which is "a lightweight, cross-platform, portable, and easy-to-maintain LaTeX distribution based on TeX Live". You can do that using the [`tinytex` package](https://cran.r-project.org/web/packages/tinytex/index.html). It also takes care of installing missing $\LaTeX$ packages on the fly when knitting documents.

```{r tinytex, eval = F}
install.packages('tinytex')

tinytex::install_tinytex()
```

*Note*: You can also install $\LaTeX$ packages manually using the `tlmgr_install()` function from the `tinytex` package.

---

## `YAML` options for PDF documents with `R Markdown`

As we have seen in the previous `R Markdown` session, there are quite a few things that you can specify in the `YAML` header to customize your output file. While most of the basic options are the same for `HTML` and PDF output (such as `title`, `author`, `toc`, or `highlight`), there also are [specific options for PDF documents](https://bookdown.org/yihui/rmarkdown/pdf-document.html). We will go through some of those in the following.

---

## Using (additional) $\LaTeX$ packages

You can also [make use of additional $\LaTeX$ packages in your `R Markdown` documents that generate PDF output](https://bookdown.org/yihui/rmarkdown-cookbook/latex-extra.html). There are different ways of including them.

As of `R Markdown` version 1.4 you can use the `extra_dependencies` option within the `output: pdf_document` setting in the `YAML` header.

```{r extra-pkg, eval=FALSE}
---
title: "It depends"
subtitle: "Contemplations on the favorite answer of scientists"
author: "The Dependables"
date: "`r Sys.Date()`"
output:
  pdf_document:
    extra_dependencies: ["threeparttable", "csquotes"]
---
```

*Note*: If you use `tinytex`, it will automatically install missing $\LaTeX$ packages on the fly when knitting documents.

---

## Using (additional) $\LaTeX$ packages

An alternative for including additional $\LaTeX$ packages (that also works with older versions of `R Markdown`) is the `header-includes` setting that takes $\LaTeX$ code (e.g., for loading packages) as a value.

```{r header-includes, eval=FALSE}
---
title: "It's all in your header"
subtitle: "What to include?"
author: "Hunter Header"
date: "`r Sys.Date()`"
header-includes:
  - \usepackage{csquotes}
output: pdf_document
---
```

*Note*: With both ways of loading additional $\LaTeX$ packages it is also possible to [specify options for loading the packages](https://bookdown.org/yihui/rmarkdown-cookbook/latex-extra.html#loading-latex-packages).

---

## Includes

To further customize and style the output, you can also include `.tex`, [`.sty` or `.cls` files](https://tug.org/pracjourn/2005-3/asknelly/nelly-sty-&-cls.pdf)<sup>1</sup> in different places in your document.

```{r includes, eval=FALSE}
---
title: "Freestyler"
subtitle: "Straight from the top of my dome"
author: "M.C. Bomfunk"
date: "`r Sys.Date()`"
output:
  pdf_document:
    includes:
      in_header: mystyle.sty
      before_body: doc-prefix.tex
      after_body: doc-suffix.tex
---
```

*Note*: Those files should be stored in the same location as the `.Rmd` file.

.small[
.footnote[
[1] Not to be confused with `.csl` files which we will talk about in the part on citations and bibliographies.
]
]

---

## $\LaTeX$ engine

As you have seen in the *Introduction to $\LaTeX$* session, there are different engines for rendering PDF documents. By default, `pdflatex` is used for rendering PDF output in `R Markdown`. You can, however, change this (to `xelatex` or `lualatex`) via the *RStudio* menu (*Tools* -> *Global options* -> *Sweave*) or by specifying the engine in the `YAML` header.

```{r engine, eval=FALSE}
---
title: "The Chaos Engine"
subtitle: "Sources of confusion about choosing the right LaTeX engine"
author: "The Engineer"
date: "`r Sys.Date()`"
output:
  pdf_document:
    latex_engine: lualatex
---
```

---

## Keeping the `.tex` file

When you knit an `R Markdwon` document to generate PDF output, the document is first converted to a $TeX$ file (with file extension `.tex`) before it is then converted to PDF via the chosen $\LaTeX$ engine. By default, the `.tex` file is automatically deleted. If you want to keep it however (e.g., for a journal submission), you can specify this in the `YAML` header. You could then, e.g., open (and further edit) the `.tex` file with a $\LaTeX$ editor like *TeXworks* or *Overleaf*.

```{r keep-tex, eval=FALSE}
---
title: "Compiler's keepers"
subtitle: "How to channel your inner TeX"
author: "Dr. Pan"
date: "`r Sys.Date()`"
output:
  pdf_document:
    keep_tex: true
---
```

---

## Custom templates

Beyond customizing the output via including `.tex` or `.sty`, you can also fully replace the default [`Pandoc` template](https://github.com/jgm/pandoc-templates/blob/master/default.latex) by specifying a different one.

```{r tex-template, eval=FALSE}
---
title: "The Template of Doom"
subtitle: "Adventures in formatting"
author: "Henry Walton Jones Jr."
date: "`r Sys.Date()`"
output:
  pdf_document:
    template: my-amazing-template.tex
---
```

*Note*: *Overleaf* offers a quite extensive (and searchable) [collection of $\LaTeX$ templates](https://de.overleaf.com/latex/templates).

---

## Template packages

There also are several `R` packages that provide different types of templates for `R Markdown`, including templates for journal articles. Two good options for that are:

- [`papaja`](https://crsh.github.io/papaja_man/) for APA manuscripts

- [`rticles`](https://github.com/rstudio/rticles) for `R Markdown` $\LaTeX$ templates for various journals and publishers

*Note*: Once you have installed packages containing `R Markdown` templates, these should also be available via the *RStudio* menu: *File* ->  *New File* -> *R Markdown* -> *From Template*.

---

## $\LaTeX$ code in `R Markdown` documents

Of course, it is also possible to use raw $\LaTeX$ outside the `header-includes` option in the `YAML` header. Just as you can use `HTML` code in `R Markdown` documents that produce `HTML` output, you can use $\LaTeX$ code in the text body of documents that generate PDF output.

For example, if you want to [insert a line break (or add an empty line) or a page break](https://de.overleaf.com/learn/latex/Line_breaks_and_blank_spaces), you can simply use the $\LaTeX$ commands `\newline` and `\newpage`.

---

## Mathematical expressions

As you may know (or remember from the previous session), $\LaTeX$ is great for typesetting mathematical expressions. There are two modes for mathematical expressions in $\LaTeX$ that you can also use in `R Markdown`: [inline and display mode](https://de.overleaf.com/learn/latex/Mathematical_expressions). Inline math expressions appear in the text, whereas expressions in display mode will be shown centered on a separate line.

For example, we could report that the mean for a variable was calculated using the formula $\bar{x} = \frac{1}{n} \sum_{i=1}^{n}x_{i}$ or present the formula for calculating the variance 
$$\sigma^{2} = \frac{\sum_{i=1}^{n}\left(x_{i} - \bar{x}\right)^{2}}{n-1}$$
*Note*: For generating formulas you can, e.g., use the [Online LaTeX Equation Editor](https://www.codecogs.com/latex/eqneditor.php) or the [Visual Math Editor](http://visualmatheditor.equatheque.net/VisualMathEditor.html).

---

## Mathematical expressions

.pull-left[
### Syntax

```
$\bar{x} = \frac{1}{n} 
  \sum_{i=1}^{n}x_{i}$

$$\sigma^{2} = \frac{\sum_{i=1}^{n} 
  \left(x_{i} - \bar{x}\right)^{2}}
  {n-1}$$

```
.small[
*Note*: There are also [other ways of starting and ending display math mode](https://de.overleaf.com/learn/latex/Mathematical_expressions). As always, there are [pros and cons to using the different options](https://tex.stackexchange.com/questions/503/why-is-preferable-to).
]
]

.pull-right[
### Output

$\bar{x} = \frac{1}{n} \sum_{i=1}^{n}x_{i}$

$$\sigma^{2} = \frac{\sum_{i=1}^{n} 
  \left(x_{i} - \bar{x}\right)^{2}}
  {n-1}$$
]

---

## Tables & PDF output

As we have discussed in the previous `R Markdown` session, there are quite a few packages for creating tables with `R`. Not all of them support PDF output (at least not out-of-the-box). The documentation website for the `gtsummary` package provides a helpful [tabular overview](https://www.danieldsjoberg.com/gtsummary/articles/rmarkdown.html) of the output options for some of the most popular packages for creating tables in `R`.

Some packages, however, allow generating `.tex` files or raw $\LaTeX$ output or can be converted to $\LaTeX$ using functions from other packages. This is, e.g., the case for the `gtsummary` package and the `stargazer` package.

---

## PDF tables with `stargazer`

We have previously used the `stargazer` to create tables for an `HTML` document. We can also use it to produce $\LaTeX$ tables for PDF documents. We just need to set the `type` argument to `latex` (instead of `html`). As in the `HTML` case we need to set the chunk option `results='asis'`. What this does is putting the raw $\LaTeX$ code into the `R Markdown` document which works fine as we have seen before with the math expressions.

```{r src-wrangling, echo=FALSE, message=FALSE}
library(tidyverse)

gp_covid <- read_csv("../../data/ZA5667_v1-0-0_CSV_synthetic-data.csv")
```

```{r stargazer-latex, eval=FALSE}
library(stargazer)

gp_covid %>% 
  na_if(98) %>% 
  select(hzcy044a, hzcy045a) %>% 
  as.data.frame() %>% 
  stargazer(type = "latex",
            header = FALSE)
```

*Note*: As before, you first need to run/source the data wrangling code before you can execute the above code example.

---

## PDF tables with `stargazer`

.tiny[
```{r ref.label = "stargazer-latex", echo = FALSE, results='asis', message=FALSE}
```
]

---

## PDF tables with `knitr::kable()`

Tables with the `kable()` function from the `knitr` package work just the same for PDF documents as they do for `HTML` output. As you may remember from the previous `R Markdown` session, however, the styling options for `knitr::kable()` are rather limited which is why tables produced with this function tend to look rather bland. This is where the package [`kableExtra`](https://haozhu233.github.io/kableExtra/) comes in which offers a wide range of additional formatting and styling options for tables created with `kable()`. It also offers options specifically for use with $\LaTeX$ and provides [detailed documentation](https://haozhu233.github.io/kableExtra/awesome_table_in_pdf.pdf) for this.

---

## Tweaking and styling tables

As we have seen, most packages for creating tables, such as `stargazer` offer various arguments or functions for customizing tables. There are also extension packages, such as `kableExtra` for `knitr::kable()` or `gtsummary` `gt`. If those options still do not satisfy your customization needs, you could copy and paste the $\LaTeX$ code produced by `stargazer` or other packages that offer this option into your `R Markdown` document and then manually edit the code there.<sup>1</sup> Even if you are not "fluent" in $\LaTeX$, if you look at the code, you should be able to figure out what most of its parts mean and do. Or you could have a look at many of the available tutorials on tables with $\LaTeX$ (such as [the one by *Overleaf*](https://de.overleaf.com/learn/latex/Tables)).

.small[
.footnote[
[1] Of course, you could do the same with `HTML` table code produced, e.g., by `stargazer`, if you want to produce `HTML` documents.
]
]

---

## Citations & bibliographies

**Disclaimer**: While we cover the subject of bibliographies and citations in this session that focuses on `R Markdown` and $\LaTeX$, you can, of course, also include those in `HTML` documents (and other types of output) produced with `R Markdown`. However, the format that is normally used for bibliographies in `R Markdown` is a $\LaTeX$ format, and there are some specific (additional) options for citations and bibliographies when using `R Markdown` in combination with $\LaTeX$ for PDF output.

---

## Citations & bibliographies

Through `Pandoc` it is possible to automatically generate citations and bibliographies in a variety of styles. It makes use of the [`citeproc` library](https://github.com/jgm/citeproc) for this purpose which "generates citations and bibliography formatted according to a CSL style". CSL stands for [Citation Style Language](https://citationstyles.org/) and there are many different CSL files for implementing various association, publisher, or journal citation styles. 

`citeproc` can handle [different bibliography file formats](https://pandoc.org/MANUAL.html#citations), but the most widely used format that also works best with `R Markdown` is $BibTeX$ (see https://en.wikipedia.org/wiki/BibTeX) which has the file extension `.bib`.

---

## Creating and managing `.bib` files

The easiest way to create and manage `.bib` files is to use a reference management software, such as the free and open-source [*Zotero*](https://www.zotero.org/).<sup>1</sup> 

If you do not want to use a reference manager, you could also use [*zoterobib*](https://zbib.org/) which is a free web service for creating bibliographies "on the fly" which can also be exported as a `.bib` file. Besides that, many journal website (and some personal website) as well as *Google Scholar* offer functions for downloading citations in $BibTeX$ format.

If you have not worked with `.bib` files before and want to see what the bibliographic entries in those look like, you can have a look at the two `.bib` files contained in the `solutions` folder in the workshop materials. You can open (and edit) those files with any text editor (such as [notepad++](https://notepad-plus-plus.org/downloads/) on *Windows*) and also directly in *RStudio*.

.small[
.footnote[
[1] You can also create and edit `.bib` files with other reference managers, such as [*Citavi*](https://www.citavi.com/en), [*Paperpile*](https://paperpile.com/?welcome), or [*Mendeley*](https://www.mendeley.com/).
]
]

---

### $BibTeX$ entries

$BibTeX$ entries can have many different fields and not all of them have to be specified. Below is an example of an entry. The first value that comes after the `{` is the citation key (in this case `GESISPanelTeam2020`) which you need for citing in in `R Markdown documents` (more on this in a bit).<sup>1</sup>

```{r bibentry, eval=FALSE}
@misc{GESISPanelTeam2020,
  title = {{GESIS Panel Special Survey on the Coronavirus SARS-CoV-2 Outbreak in GermanyGESIS Panel Special Survey on the Coronavirus SARS-CoV-2 Outbreak in Germany}},
  author = {{GESIS Panel Team}},
  year = {2020},
  publisher = {{GESIS Data Archive}},
  doi = {10.4232/1.13520}
  
}
```

.small[
.footnote[
[1] There are different conventions for citation keys and most reference management tools offer options for formatting them in a unified way.
]
]

---

## Using `.bib` files in `R Markdown`

You can use `.bib` files in `R Markdown` documents by listing them as values for the `bibliography` option in the `YAML` header. You can list one `.bib` file...

```{r bib-one, eval = FALSE}
---
title: "Cite it right"
subtitle: "Making reference management great again"
author: "Sir Cite-a-Lot"
date: "`r Sys.Date()`"
output:
  pdf_document
bibliography: references.bib
---
```

---

## Using `.bib` files in `R Markdown`

... or multiple ones.

```{r bib-mult, eval = FALSE}
---
title: "What a Cite!"
subtitle: "What you cite is what you get"
author: "Bob Lee Ography"
date: "`r Sys.Date()`"
output:
  pdf_document
bibliography: [references1.bib, references2.bib]
---
```

*Note*: Again, the `.bib` file(s) should be stored in the same place as the `.Rmd` file.

---

## Citation styles

When writing a paper (or some other publication), we typically want (or rather have) to adhere to specific citation styles. These can also be employed in `R Markdown` via `Pandoc` (or rather `citeproc`) by using a Citation Style Language `.csl` file and specifying it in the `YAML` header.

```{r csl, eval=FALSE}
---
title: "Citing in style"
subtitle: "A guide to making journal editors happy"
author: "Reviewer 2"
date: "`r Sys.Date()`"
output:
  pdf_document
bibliography: references.bib
csl: apa.csl
---
```

---

## Citation styles

The free and open-source reference management tool [*Zotero*](https://www.zotero.org/) offers a [searchable collection of citation styles](https://www.zotero.org/styles/) from which the associated `.csl` files can also be downloaded. Another comprehesive source is the [`citation-style-language/styles` *GitHub* repository](https://github.com/citation-style-language/styles).

---

## Citations & bibliographies with $\LaTeX$ packages

Instead of using the functionalities of `citeproc` (in combination with `.csl` files), for PDF output it is also possible to use the $\LaTeX$ packages [`natbib`](https://ctan.org/pkg/natbib) or [`biblatex`](https://ctan.org/pkg/biblatex) for including and formatting citations and bibliographies. These support different bibliography styles. You can specify both the $\LaTeX$ citation package and one of the styles it supports in the `YAML` header.

```{r latex-cite, eval=FALSE}
---
title: "Citing in style"
subtitle: "A guide to making journal editors happy"
author: "Reviewer 2"
date: "`r Sys.Date()`"
output:
  pdf_document:
  citation_package: biblatex
bibliography: references.bib
blibio-style: apa
---
```

---

## Citing references

Once you have created and specified a `.bib` file containing references, you can cite them in your `R Markdown` document using their citation keys (which are the first fields/values in the $BibTeX$ entries). 
There are different commands for different types of citations:

- `@citekey`: citation without parentheses
- `[@citekey]`: citation in parentheses
- `[@citekey1; @citekey2]`: multiple citations in parentheses
- `[-@key]`: citation in parentheses without author names (i.e., just the year)

As soon as you cite at least one reference that way in your `R Markdown` document, a bibliography will be automatically inserted at the end of your document.<sup>1</sup>

.small[
.footnote[
[1] It is, however also possible to [include items in the bibliography without citing them in the text](https://bookdown.org/yihui/rmarkdown-cookbook/bibliography.html#add-an-item-to-a-bibliography-without-using-it).
]
]

---

## Citing data & FOSS

One point we would like to mention here is that, in addition to citing (relevant) literature, it is important to properly cite any data as well as free and open-source (FOSS) software you use, such as `R` and its packages. Citations for data sets are usually provided by (curated) archives. For finding out how to cite `R` and its packages you can use the `citation()` function.

.smaller[
```{r cite-r}
citation()
```
]

---

## Citing `R` packages

```{r cite-pkg}
citation("stargazer")
```

---

## `R` packages for citations & bibliographies

As you might have guessed, there also are `R` packages that can facilitate working with citations and bibliographies in `R Markdown`.

A really easy-to-use option is the [`citr` package](https://github.com/crsh/citr) which provides a handy *RStudio* addin for automatically inserting citations and bibliographies in `R Markdown` documents and also allows direct connection to existing *Zotero* libraries.

The [`bbt` package](https://github.com/paleolimbot/rbbt) also provides an *RStudio* addin and *Zotero* integration, including support for [*Better BibTeX* *Zotero* extension](https://retorque.re/zotero-better-bibtex/).

Another package that can facilitate handling bibliographies in `R Markdown` is [`RefManageR`](https://github.com/ropensci/RefManageR/). 

The [`knitcitations`](https://github.com/cboettig/knitcitations) package can be used to directly cite by DOI or URL (instead of by citation key) in a `R Markdown` document.

---

## Outlook: Even more advanced use of `R Markdown`

An even more advanced use of `R Markdown` is the production of parameterized reports. This can, e.g., be used to produce similar reports for different data sets or different groups within a data set. Explaining this would be well beyond the scope of this workshop, but there are some good tutorials on this available online (e.g., in the [*RStudio* documentation](https://docs.rstudio.com/connect/1.7.4/user/param-rmarkdown.html) and in chapters in the books [*R Markdown: The Definitive Guide*](https://bookdown.org/yihui/rmarkdown/parameterized-reports.html) and the [*R Markdown Cookbook*](https://bookdown.org/yihui/rmarkdown-cookbook/parameterized-reports.html)).

A good example of what can be done with parameterized reports in `R Markdown` are the [wave reports of the *GESIS Panel*](https://www.gesis.org/en/gesis-panel/documentation).

---

class: center, middle

# [Exercise](https://jobreu.github.io/reproducible-research-gesis-2021/exercises/B2_RMarkdown-LaTeX_exercises_question.html) time `r emo::ji("weight_lifting_woman")``r emo::ji("muscle")``r emo::ji("running_man")``r emo::ji("biking_man")`

## [Solutions](https://jobreu.github.io/reproducible-research-gesis-2021/solutions/B2_RMarkdown-LaTeX_exercises_solution.html)
